%%-*- mode: LaTeX; mode: FlySpell; -*-

\documentclass{elsarticle}

\input{preamble}

\begin{document}

\title{Blur And Collapse}


\author[ncl]{P. ~Missier\corref{cor1}}
\ead{Paolo.Missier@newcastle.ac.uk}

\author[cov]{J. ~Bryans\corref{cor2}}
\ead{Jeremy.Bryans@coventry.ac.uk}


\address[ncl]{School of Computing Science, Newcastle University, UK}
\address[cov]{Institute for Future Transport and Cities, Coventry University, UK}

\cortext[cor1]{Principal Corresponding Author}
\cortext[cor2]{Corresponding Author}

%\markboth{IEEE Transaction on Knowledge and Data Engineering,~Vol.~X, No.~Y, DATE}%
%{Shell \MakeLowercase{\textit{et al.}}: happy}


\begin{abstract}

\end{abstract}

\begin{keyword}
Provenance \sep Provenance metadata \sep provenance abstraction 
\end{keyword}

\maketitle



\section{Introduction and motivation}

\mnote{Running examples:  we may need two, one for each operator.
	
	PM: suggest picking p-graphs that are plausible when viewed as the trace of a program, ie using abstract function names for activities.
}
	
\mnote{main contributions:
	\begin{itemize}
	\item three graph rewriting operators for abstracting over provenance graphs, \jwb{blur (two versions of blur, collapse, sibling collapse)}
          formalised and with proof that the rewriting preserves PROV validity
		\item algebraic properties of the operators (commutative??)
		\item small PROV extension to introduce new relationships when the existing ones are generalised i.e. to ``influence'', that is the new relationship types qualify ``influence''
               \item use Janetta's tool to show these available in an implementation
                  \item (AMBITION) demonstrate these useful as comprehension tools (so not obfuscation..) to the urban observatory as alternate interfaces to theor data for the general public and experts. 
%		\item demonstration of how the operators support a variety of abstraction policies and how we cna use a variety of graph metrics to measure the effect of applying a specific policy to a graph
%		\item empirical evaluation of the rewriting i.e. using (1) a variety of synthetic graphs, generated using ProvGen, (2) a variety of policies chosen as representative examples and (3) a variety of provenance network metrics
 	\end{itemize}
	
}
	
	
\section{The Blur and Collapse operators}
	
	
	\mnote{HOW ABOUT CUTTING RELATIONSHIPS}
	
	\subsection{Formal definitions}
\mnote{we decided:
	
	\begin{itemize}
		\item three/four separate formal definitions, one for each op.
		\item blur does not follow from collapse
		\item blur +/- does not imply blur (CHECK)
		\item collapse limited to a-a  (supported by intuition on what it means to collapse. collapsing e-e means you merge inputs and outputs which does not make much sense)
  		\item collapse requires using the PROV top-level ``influence'' relationships (only between two activities) --> show an example. This is well supported by intuition: when $a$ uses the intermediate data produced by $a_1$ and consumed by $a_2$, and we collapse $a_1, a_2$ into $a'$, then $a$ becomes related to $a'$. There influence has a natural interpretation.
  		  \item \jwb{siblings is a kind of ``horizontal'' collapse.}
	\end{itemize}
}

\subsection{Initial definitions}

\jwb{begin with definition of graphs, and any others we need from previous paper: in(v), out(v) and in and out of a set of nodes. }
\jwb{also definitions of $\wgby$ and $\used$ and $\infl$}

If it exists in a graph $G$, the $\Path$ from $x$ to $y$ is given by $\Path(x,y,G)$. The \emph{length} of the path from $x$ to $y$ is given by $\len(\Path(x,y,G))$, and is defined as the number of edges between the $x$ and $y$. 

\subsection{Collapse}

We give now the definition of the collapse operator $\col$, adjusted to take  all possible paths into account.
%
Informally, the collapse operator takes a graph, and two activities at different temporal points in that graph (is it right to say one logically precedes the other?) and removes both them and all paths between them, replacing all with a single new activity. 


\begin{definition}[$\col$]  \label{def:col}
\jwb{note we've agreed that collapse should be a-a only, but that's not built into this definition yet}.  For nodes $x$, $y$ in a graph $G = (V,E)$ such that $x \le y$, and a new node $v_{N1}$ not in $V$,  $\col(x,y,G)$ is defined provided there is a path between nodes $x$ and $y$. Let $S$ be the set of all nodes on all paths from $x$ to $y$.
  %
  Formally, $S=\Union \{s|x \le s \land s \le y\}$.
  %
  Let $r(l)$ stand for the function $\relabel(l)$. Then $\col(x,y,G) =  G'$, such that $G'=(V',E')$, where 
  \begin{eqnarray*}
  V' & = & (V\hide S) \union \{v_{N1}\}     \\
  E' & = &  E\hide (
                   \{(v',v,l)| (v',v,l) \in E, v',v \in S\}
%                   \union
%                   \{(v,v',l)| v,v' \in P\}
                  )\\
  && \union\ \{(v',v_{N1},r(l))|(v',v,l) \in E, v\in S, v'\nin S\}\\
                   && \union\ \{(v_{N1},v',r(l))|(v,v',l) \in E, v\in S, v' \nin S\}\\
  \end{eqnarray*}
\end{definition}
If $type(v)$ returns the type of a node $v$, the function $relabel$ (abbreviated above as $r$) is defined as 
\begin{definition}[$relabel$] \label{def:relabel}
  For an edge $(v,v',l)$, 
  \[
   (v,v',relabel(l)) = \left\{
   \begin{array}{l}
      l,    {\emph{if}}\; type(v) \neq type(v') \\
      infl, \emph{otherwise}
   \end{array}   \right.
  \]
\end{definition}
\noindent
$relabel$ is promoted in the obvious way to sets of edges.

\subsection{Variations on blur}

\jwb{have to retain all values (in all levels?) as part of the definition of the blurs, in order that the operators can be reversed}

The $\blur$ operator has two versions: directed and non-directed. The definition for directed blur uses the definition for a directed blurset, given below.

\begin{definition}[{\it Directed} $\blurset$] \label{def:directed-blurset}
  For $n \in \nat$,
  \begin{eqnarray*}
    \dblurset(n,v,G) & = & \{v'|\len(\Path(v,v',G)) \leq n\}\union \\
                     &   & \{v'|\len(\Path(v',v,G)) \leq n\}.
  \end{eqnarray*}
\end{definition}


\begin{definition}[{\it Directed} $\blur$] \label{def:directed-blur}
  If $G = (V,E)$ with $v \in V$, and $n_b$ is a new node (\jwb{of type?}) then $\dblur(n,v,G) = (V',E')$, where
  \begin{eqnarray*}
  V' &=& (V\hide \dblurset(n,v,G)) \union\ \{n_b\}\\
  E' &=& E\hide \{(v_1,v_2,l)| v_1 \in \dblurset(n,v,G) ~\emph{or}~ v_2 \in \dblurset(n,v,G)\}\\
  &&\union\ \{(n_b,v',l)|(v_1,v',l) \in E ~\emph{and}~ v_1 \in \dblurset(n,v,G) ~\emph{and}~ v' \nin \dblurset(n,v,G) \}\\
  &&\union\ \{(v',n_b,l)|(v',v_1,l) \in E ~\emph{and}~ v_1 \in \dblurset(n,v,G)  ~\emph{and}~ v' \nin \dblurset(n,v,G) \} \\
  \end{eqnarray*}
\end{definition}

Undirected blur includes nodes which are linked, but do not lie on a direct path. 
%
To define this, we begin with by extracting an undirected graph, in which the directional information has been removed from the edges, from the directed one. In a direct graph $(V,E)$, an edge $(v_1,v_2,l)$ is an edge beginning at $v_1$ and ending at $v_2$ (with the label $l$). In the undirected version of the graph labelling is removed, and all edges are replaced by undirected edges, so $(v_1,v_2)$ is now a non-directed edge between $v_1$ and $v_2$ and has no directionality. 

%If  $v_1$ and $v_2$ are connected in the undirected graph, we define a $



\begin{definition}[{\it Undirected} $\blurset$] \label{def:undirected-blurset}
  If $uG$ is the undirected version of a graph $G$, $v \in G$, and $n \in \nat$,
  \begin{eqnarray*}
    \udblurset(n,v,G) & = & \{v'|\len(\Path(v,v',uG)) \leq n\}\union \\
                     &   & \{v'|\len(\Path(v',v,uG)) \leq n\}.
  \end{eqnarray*}
\end{definition}

\begin{definition}[{\it Undirected} $\blur$] \label{def:undirected-blur}
  If $G = (V,E)$ with $v \in V$, and $n_b$ is a new node (\jwb{of type?}), then $\udblur(n,v,G) = (V',E')$, where
  \begin{eqnarray*}
  V' &=& (V\hide \udblurset(n,v,G)) \union\ \{n_b\}\\
  E' &=& E\hide \{(v_1,v_2,l)| v_1 \in \udblurset(n,v,G) ~\emph{or}~ v_2 \in \udblurset(n,v,G)\}\\
  &&\union\ \{(n_b,v',l)|(v_1,v',l) \in E ~\emph{and}~ v_1 \in \udblurset(n,v,G)  ~\emph{and} v' \nin \udblurset(n,v,G) \} \\
  &&\union\ \{(v',n_b,l)|(v',v_1,l) \in E ~\emph{and}~ v_1 \in \udblurset(n,v,G)  ~\emph{and} v' \nin \udblurset(n,v,G) \} \\
  \end{eqnarray*}
\end{definition}



\subsection{Sibling collapse}

This set of operators collapse siblings into a single node. We begin with some definitions.

The \emph{source} of a node is the set of all nodes which are directly linked via a single edge into the root node.  Recall providence graphs point into the past. Sources are defined using the definition of $out$ edges. 

\begin{definition}[$\source$] \label{def:source}
  Given a graph $G = (V,E)$ and a node $v \in V$,
  \[
  \source(v) = \Union\{v_s|(v,v_s,l) \in out(v)\}.
  \]
\end{definition}

Conversely, the $\sink$ of a node is all nodes which are the end points of edges from the root node.  It is defined using the definition of $in$ edges.

\begin{definition}[$\sink$] \label{def:sink}
  Given a graph $G = (V,E)$ and a node $v \in V$,
  \[
  \source(v) = \Union\{v_s|(v_s,v,l) \in in(v)\}.
  \]
\end{definition}

Two nodes are $\fullsiblings$ if their source sets and sink sets are all identical.  

\begin{definition}[$\fullsiblings$] \label{def:fullsiblings}
  Given a graph $G = (V,E)$ and two nodes $v,w \in V$, $v$ and $w$ are $\fullsiblings$ if $\source(v) = \source(w)$ and $\sink(v) = \sink(w)$. 
\end{definition}

Two nodes are $\halfsiblings$ if their source sets (sink sets) share a common element.  

\begin{definition}[$\halfsiblings$] \label{def:fullsiblings}
  Given a graph $G = (V,E)$ and two nodes $v,w \in V$, $v$ and $w$ are $\halfsiblings$ if $\source(v) \inter \source(w) \ne \emptyset$ and $\sink(v) \inter \sink(w) \ne \emptyset$. 
\end{definition}

These definitions can be extended to more than two siblings. 

\jwb{Sibling abstraction is defined as:}

\begin{definition}
  Given a graph $G = (V,E)$ and two nodes $v,w \in V$, $v$ and $w$ are $\halfsiblings$ i
\end{definition}



\subsection{Composing Blur operations}
  
  	\begin{itemize}
	\item \jwb{compositionality,  commutativity don't apply; not binary operators}
	\item thus we define a explicit semantics for dealing with a set of Blur operations -  \textit{union semantics}
\end{itemize}


  \subsection{Composing Collapse operations}

\mnote{  
  	\begin{itemize}
  	  \item collapse is compositional, probably commutative  --> \textbf{proof??}
  	  \item if this is the case, then no need for union semantics
  	\end{itemize}
  	}
  	
  \subsection{Composing Blur and Collapse operations}  
		
	\mnote\textbf{{?? not worked out.} }

\section{Abstraction policies}

\mnote{We want to show how different families of abstraction policies can be specified and enforced using these two operators and their composition.}

\subsection{Policy definition}

\mnote{What is the most abstract definition of an abstraction policy?
	
it consists of two elements:

  	\begin{itemize}
	\item specifying the nodes that are arguments to the operators (and the radius in the case of blur). This specification is either explicit (mentioning specific nodes) or intensional, ie through conditions that predicate on properties of the nodes, eg sensitivity, utility.
	
	\item apply the set of operations. If collapse only then probably commutative (order irrelevant), blur: defined by union semantics for blur. if a mix: not clear.
	
  \end{itemize}
}

\subsection{Policy Examples}
	
		\mnote{1. sensitivity threshold on single node (see old paper). 2. average sensitivity within a set of neighbouring nodes.. 3....}
		
	\mnote{apply each policy to our running example and show the resulting graph}

\section{Empirical evaluation}

\mnote{the aim of the evaluation is to show how the framework allows for the easy specification of policies and comparing their effects on a collection of test graphs, using a variety of provenance network metrics}

\mnote{introduce network metrics: sensitivity, utility, plus provenance metrics as proposed by Moreau 2018}

\mnote{experimental testbed consists of a set of PROV graphs generated using provGen}

\mnote{apply each policy to each graph, measure metrics before /  after, compare as charts / tables}

\section*{Acknowledgements}

The support of the ONRG and the EPSRC in funding this research is gratefully acknowledged. 

\appendix

\section{Proofs}

\input{proof}




\section*{References}

%\bibliographystyle{plain}
%\bibliography{prov-abstraction-foundations,p3s-JB}

%\appendix


\end{document}
